---
title: "Crime Rates in the US"
author: "Jade Towgood"
date: "`r Sys.Date()`"
output: ioslides_presentation
mathjax: local
---
<style type="text/css">
body p, div, h1, h2, h3, h4, h5 {
color: black;
font-family: Modern Computer Roman;
}
slides > slide.title-slide hgroup h1 {
color: #8C1D40;
}
h2 {
color: #8C1D40; 
}
</style> 

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(datasets)
library(tinytex)
library(plotly)
library(ggplot2)
library(dplyr)
```

```{r}
data("USArrests")
```
## Data Set `USArrests`
This data set displays the violent crime rates per 100,000 residents by US State in 1973. <br>
<br>
There are 4 variables with 50 states in this data set: <br>

  - murder <br>
  - assault <br>
  - rape <br>
  - urban population

## The Question
Is there a correlation between the violent crime rates and each other? <br>

  - rape vs murder <br>
  - murder vs assault <br>
  - assault vs rape <br>

## Summary Violent Crimes: Murder
```{r, echo=FALSE}
boxplot(USArrests$Murder, horizontal = TRUE, xlab = "Rate (cases/100,000 people)", main = "Murder Crimes", col = "red")
summary(USArrests$Murder)
```

## Summary Violent Crimes: Assault
```{r, echo=FALSE}
boxplot(USArrests$Assault, horizontal = TRUE, xlab = "Rate (cases/100,000 people)", main = "Assault Crimes", col = "orange")
summary(USArrests$Assault)
```

## Summary Violent Crimes: Rape
```{r, echo=FALSE}
boxplot(USArrests$Rape, horizontal = TRUE, xlab = "Rate (cases/100,000 people)", main = "Rape Crimes", col = "gold")
summary(USArrests$Rape)
```
## Explanation of Code: Boxplots
```{r, eval=FALSE, echo=TRUE}
boxplot(USArrests$Murder, horizontal = TRUE, xlab = "Rate 
        (cases/100,000 people)", main = "Murder Crimes", col = "red") 
summary(USArrests$Murder) 

boxplot(USArrests$Assault, horizontal = TRUE, xlab = "Rate 
        (cases/100,000 people)", main = "Assault Crimes", col = 
          "orange") 
summary(USArrests$Assault) 

boxplot(USArrests$Rape, horizontal = TRUE, xlab = "Rate 
        (cases/100,000 people)", main = "Rape Crimes", col = "gold") 
summary(USArrests$Rape) 
```
I decided to do 3 separate boxplots and summaries to make the data easier
to read and understand it instead of putting it all together. I excluded 
the urban population data because it was not a variable I was discussing.

## Scatter Plots: Rape vs Murder
```{r}
murderRates <- USArrests$Murder
rapeRates <- USArrests$Rape

x = rapeRates
y = murderRates

plot_ly(x = x, y = y, type = "scatter", mode = "markers") %>%
  add_trace(x = rapeRates, y = predict(lm(murderRates~rapeRates)), mode = "lines", type = "scatter", line = list(color = "Red")) %>%
        layout(title = "Correlation between Murder and Rape Crimes in the US", xaxis = list(title = "Rape Rates"), yaxis = list(title = "Murder Rates"))
```
This scatter plot displays a positive correlation between the rate of rape and murder crimes.

## Scatter Plots: Assault vs Murder
```{r}
assultRates <- USArrests$Assault
murderRates <- USArrests$Murder

x = assultRates
y = murderRates

plot_ly(x = x, y = y, type = "scatter", mode = "markers") %>%
  add_trace(x = assultRates, y = predict(lm(murderRates~assultRates)), mode = "lines", type = "scatter", line = list(color = "Red")) %>%
        layout(title = "Correlation between Assault and Murder Crimes in the US", xaxis = list(title = "Assault Rates"), yaxis = list(title = "Murder Rates"))
```
This scatter plot displays a positive correlation between the rate of assault and murder crimes.

## Scatter Plots: Assault vs Rape
```{r}
assultRates <- USArrests$Assault
rapeRates <- USArrests$Rape

x = assultRates
y = rapeRates

plot_ly(x = x, y = y, type = "scatter", mode = "markers") %>%
  add_trace(x = assultRates, y = predict(lm(rapeRates~assultRates)), mode = "lines", type = "scatter", line = list(color = "Red")) %>%
        layout(title = "Correlation between Assault and Rape Crimes in the US", xaxis = list(title = "Assault Rates"), yaxis = list(title = "Rape Rates"))
```
This scatter plot displays a positive correlation between the rate of assault and rape crimes.

## Explanation of Code: Scatterplots
```{r, eval=FALSE, echo=TRUE}
murderRates <- USArrests$Murder
rapeRates <- USArrests$Rape

x = rapeRates
y = murderRates

plot_ly(x = x, y = y, type = "scatter", mode = "markers") %>%
  add_trace(x = rapeRates, y = predict(lm(murderRates~rapeRates)), 
            mode = "lines", type = "scatter", line = list(color = 
            "Red")) %>%
        layout(title = "Correlation between Murder and Rape Crimes in 
        the US", xaxis = list(title = "Rape Rates"), yaxis = list
          (title = "Murder Rates"))
```

## Cont.
```{r, eval=FALSE, echo=TRUE}
assultRates <- USArrests$Assault
murderRates <- USArrests$Murder

x = assultRates
y = murderRates

plot_ly(x = x, y = y, type = "scatter", mode = "markers") %>%
  add_trace(x = assultRates, y = predict(lm(murderRates~assultRates)), 
            mode = "lines", type = "scatter", line = list(color = 
            "Red")) %>%
        layout(title = "Correlation between Assault and Murder Crimes 
        in the US", xaxis = list(title = "Assault Rates"), yaxis = 
          list(title = "Murder Rates"))
```

## Cont.
```{r, eval=FALSE, echo=TRUE}
assultRates <- USArrests$Assault
rapeRates <- USArrests$Rape

x = assultRates
y = rapeRates

plot_ly(x = x, y = y, type = "scatter", mode = "markers") %>%
  add_trace(x = assultRates, y = predict(lm(rapeRates~assultRates)), 
            mode = "lines", type = "scatter", line = list(color = 
            "Red")) %>%
        layout(title = "Correlation between Assault and Rape Crimes 
        in the US", xaxis = list(title = "Assault Rates"), yaxis = 
          list(title = "Rape Rates"))
```
I decided that scatterplots would be the best way to analyze whether there
was a correlation with a linear regression model included. 

## Linear Regression Model: \(R^2\) Value
```{r, echo=FALSE}
assultRates <- USArrests$Assault
rapeRates <- USArrests$Rape
murderRates <- USArrests$Murder

x1 = assultRates
y1 = rapeRates
y2 = murderRates
```
Rape vs Murder
```{r}
lm_model <- lm(y2~y1, data = USArrests)
r_squared <- summary(lm_model)$r.squared
cat("R-squared value:", r_squared, "\n")
```
Assault vs Murder
```{r}
lm_model <- lm(y2~x1, data = USArrests)
r_squared <- summary(lm_model)$r.squared
cat("R-squared value:", r_squared, "\n")
```
Assault vs Rape
```{r}
lm_model <- lm(y1~x1, data = USArrests)
r_squared <- summary(lm_model)$r.squared
cat("R-squared value:", r_squared, "\n")
```

## Explanation of Code: Linear Regression Model
```{r, eval=FALSE, echo=TRUE}
assultRates <- USArrests$Assault
rapeRates <- USArrests$Rape
murderRates <- USArrests$Murder

x1 = assultRates
y1 = rapeRates
y2 = murderRates

lm_model <- lm(y2~y1, data = USArrests)
r_squared <- summary(lm_model)$r.squared
cat("R-squared value:", r_squared, "\n")
```


## Conclusion
Assuming: <br>

- The size of error doesn't change significantly across the values. <br>
- There are no hidden relationships among the variables. <br>
- The data has a normal distribution. <br>
- The Relationship between the variables are linear. <br>
<br>

There seems to be a positive correlation between all crimes in this data set, which answers my question of whether each violent crime has some type of correlation with one another.